
%% Thesis information:
\title{On the Quantum Base Size of a Group}
\author{Tristan Wylde-LaRue}
% The month and year that you submit your FINAL draft TO THE LIBRARY (May or December)
\date{May 2019}
\division{Mathematics and Natural Sciences}
\advisor{Jamie Pommersheim}
%If you have two advisors for some reason, you can use the following
%\altadvisor{Your Other Advisor}
%%% Remember to use the correct department!
\department{Mathematics}






  \maketitle
  \frontmatter % this stuff will be roman-numbered
  \pagestyle{empty} % this removes page numbers from the frontmatter

% Acknowledgements (Acceptable American spelling) are optional
% So are Acknowledgments (proper English spelling)
    \chapter*{Acknowledgements}
    First and foremost, I would like to thank my family for their incredible patience and kindness, I could not 
    have made it through this without them. I also need to express my overwhelming thanks to Roger for the 
    motivation and support he gave me.

    I am extremely grateful for my advisor Jamie who guided me through all of this. It was amazing to have been 
    able to take classes with him and share in his excitement for everything math related. Undoubtedly, he is one 
    of the best professors I have ever had. For thesis, he has been nothing but kind and patient. It's been a 
    pleasure getting to work with him individually over the past year. It's very sad to think that I won't ever 
    have another Jamie class.

    The entire math community at Reed has been nothing but great to me in my time here. I have to thank all of my 
    friends in the math department for the long nights working on real analysis problem sets. Without them, I 
    certainly would not have enjoyed my math education as much as I did.

    Finally, I want to thank all of my friends who have put up with me this last year working on this. I wish I 
    could name everyone individually, but it would be an extremely long list. However, I need to give special 
    thanks to Raina, Duncan and Sarina for really getting me through this year. Parts of this year have been very 
    rough and they were always there to help without any second thoughts. I'm infinitely indebted to them.


% The preface is optional
% To remove it, comment it out or delete it.
    %\chapter*{Preface}
	%This is an example of a thesis setup to use the reed thesis document class.
	
	

    %\chapter*{List of Abbreviations}
		%You can always change the way your abbreviations are formatted. Play around with it yourself, use tables, or come to CUS if you'd like to change the way it looks. You can also completely remove this chapter if you have no need for a list of abbreviations. Here is an example of what this could look like:

	%\begin{table}[h]
	%\centering % You could remove this to move table to the left
	%\begin{tabular}{ll}
		%\textbf{ABC}  	&  American Broadcasting Company \\
		%\textbf{CBS}  	&  Columbia Broadcasting System\\
		%\textbf{CDC}  	&  Center for Disease Control \\
		%\textbf{CIA}  	&  Central Intelligence Agency\\
		%\textbf{CLBR} 	&  Center for Life Beyond Reed\\
		%\textbf{CUS}  	&  Computer User Services\\
		%\textbf{FBI}  	&  Federal Bureau of Investigation\\
		%\textbf{NBC}  	&  National Broadcasting Corporation\\
	%\end{tabular}
	%\end{table}
	

    \tableofcontents
% if you want a list of tables, optional
    %\listoftables
% if you want a list of figures, also optional
    %\listoffigures

% The abstract is not required if you're writing a creative thesis (but aren't they all?)
% If your abstract is longer than a page, there may be a formatting issue.
    \chapter*{Abstract}
    In this thesis we study the query complexity of quantum oracle problems in which a learner must identify a 
    hidden group element by querying its action on a set. A recent paper of Pommersheim and Copeland connects this 
    problem to a classical problem in the representation theory of finite groups. We extend their work by computing 
    the query complexity for symmetric oracle problems associated to the dihedral groups.
    
    \chapter*{Dedication}
    \emph{To my sister Cordelia who inspired my love of math}.
  \mainmatter % here the regular arabic numbering starts
  \pagestyle{fancyplain} % turns page numbering back on

