

\chapter{Constructions of the Quantum Base Size}

This chapter gives explicit computations of the \emph{quantum base size} defined in Chapter 3. We prove the number 
of queries required for exact and bounded error learning for the dihedral groups $D_{4k}$ and the general linear 
groups $\GL(2, q)$.

We work through the example of the dihedral groups in much detail to illustrate how one might go about computing 
these bounds for more families of groups.





\section{Character Table of the Dihedral Group}

The \emph{dihedral group} of order $2n$ is given by
\[
    D_{2n} := \langle a,b \mid a^n = b^2 = 1,\ bab = a^{-1} \rangle
\]
It arises naturally as the planar symmetries of the $n$-gon. For this reason, elements of the form $a^i$ are often 
referred to as rotations and elements of the form $a^ib$ are referred to as reflections.

The character table for the case when $n$ is even is slightly different from the case when $n$ is odd. To keep 
things as clear as possible and to avoid too much repetition, we will do everything only for the case where $n$ is 
even. The results for the odd case readily follow from the exact same approach, only there are fewer conjugacy 
classes and thus fewer cases to check.

\subsection{Conjugacy Classes of $D_{4k}$}

Another writeup of this can be found in chapter 12 of \cite{James&Liebeck}
The conjugacy classes will be slightly different if $n$ is even or odd. We first give the case for $n$ even. 

Suppose $n = 2k$ (we are considering the Dihedral group of order $4k$)

The center of $D_8$ is given by
\[
    \mathbf{Z}(D_8) = \{1, a^\}
\]
Thus $[1]$ and $[a^k]$ are both of size 1.


Let $a^i$ be a rotation. Then $ba^ib^{-1} = a^{-1}$



\subsection{Irreducible Characters of $D_8$}

Again, this will be different depending on whether $n$ is even or odd. 

First we give the even case, for $D_{4k}$.

\subsubsection{Linear Characters}
The linear characters are precisely the homomorphisms $\lambda : D_{4k} \rightarrow \C^\times$. We can compute all 
such homomorphisms by looking at how they must act on the generators. 

Let $\lambda : D_{4k} \rightarrow \C^\times$ be a homomorphism. Since $b^2 = 1$ we must have $\lambda(b)^2 = 1$, 
hence the only possible values for $\lambda(b)$ are $1$ and $-1$. Since we have the relation $abab = 1$, then $1 = 
\lambda(abab) = \lambda(a)\lambda(b)\lambda(a)\lambda(b) = \lambda(a)^2$. Thus $\lambda(a)$ must also be either $1$ 
or $-1$. This leaves us with four possible choices, each of which is a linear character of $D_{4k}$

The following table gives the four linear characters of $D_{4k}$
\begin{center}
\begin{tabular}{c|ccccc}
    $g$ &$1$ & $a^k$ & $a^i$ & $b$ & $ab$ \\ \hline
    $\lambda_1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
    $\lambda_2$ & $1$ & $1$ & $1$ & $-1$ & $-1$ \\
    $\lambda_3$ & $1$ & $(-1)^k$ & $(-1)^i$ & $1$ & $-1$ \\
    $\lambda_4$ & $1$ & $(-1)^k$ & $(-1)^i$ & $-1$ & $1$ \\
\end{tabular}
\end{center}

\subsubsection{The Other Irreducible Characters}
Consider the family of representations $\{\rho_j\}$ given by parameter $1 \leq j \leq k-1$ defined by
\[
    a \mapsto \rmat{\zeta_{2z}^j}{0}{0}{\zeta_{2k}^{-j}}, \quad \ \ b \mapsto \rmat{0}{1}{1}{0}
\]

Let $\psi_j$ denote the character of the representation $\rho_j$. Taking the trace gives us
\begin{align*}
    &\psi_j(a^nb^m) = \begin{cases}
        \zeta_n^j + \zeta_n^{-j}, &\text{if } b = 0 \\
        0,                &\text{if } b =1
                    \end{cases}
\end{align*}
In terms of the conjugacy classes this is
\begin{center}
\begin{tabular}{c|ccccc}
    $g$ &$1$ & $a^k$ & $a^i$ & $b$ & $ab$ \\ \hline
    $\psi_j$ & $ \ \ 2 \ \ $ & $\ 2(-1)^j\ $ & $\zeta_{2k}^{ij} + \zeta_{2k}^{-ij}$ & $\ \ 0 \ \ $ & $ \ \ 0 \ \ $
\end{tabular}
\end{center}

In order to show $\psi_j$ is irreducible we compute $\langle \psi_j, \psi_j \rangle$.
\begin{align*}
    \langle \psi_j, \psi_j \rangle &= \frac{1}{|G|}\sum_{g \in G} \psi_j(g) \overline{\psi_j(g)} \\
                           &= \frac{1}{4k}\left[2^2 + \bigl(2(-1)^j\bigr)^2 + 
                               2\sum_{i=1}^{k-1}\left(\zeta_{2k}^{ij} + 
                               \zeta_{2k}^{-ij}\right)\left(\overline{\zeta_{2k}^{ij} + 
                       \zeta_{2k}^{-ij}}\right)\right] \\
%                             
               &= \frac{1}{4k}\left[ 4 + 4 + 2\sum_{i=1}^{k-1}\left(\zeta_{2k}^{ij} + \zeta_{2k}^{-ij}\right)^2 
               \right] \\
               &= \frac{1}{4k}(8 + 2(4)(k-1)) \\
               &= 4k \end{align*}

Thus by Theorem 2.7, $\psi_j$ is an irreducible character. 

We have found 4 irreducible linear characters and $k-1$ irreducible characters of degree 2. We have shown $D_{4k}$ 
has $k+3$ conjugacy classes and so this must be a complete set of characters. Hence the character table of $D_{4k}$ 
is 

\begin{center}
\begin{tabular}{c|ccccc}
    $g$ &$1$ & $a^k$ & $a^i$ & $b$ & $ab$ \\ \hline
    $\lambda_1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
    $\lambda_2$ & $1$ & $1$ & $1$ & $-1$ & $-1$ \\
    $\lambda_3$ & $1$ & $(-1)^k$ & $(-1)^i$ & $1$ & $-1$ \\
    $\lambda_4$ & $1$ & $(-1)^k$ & $(-1)^i$ & $-1$ & $1$ \\
    $\psi_j$ & $2$ & $2(-1)^j$ & $\zeta_{2k}^{ij} + \zeta_{2k}^{-ij}$ & $0$ & $0$
\end{tabular}
\end{center}

\section{Quantum Base Size of the Dihedral Group}

Here we compute the quantum base size for the standard action of the Dihedral group of order $2n$ on the vertices 
of the $n$-gon.


We will go through the even case in detail and then summarize the results for the odd case.




\subsection{Permutation Character}



Let $G = D_{4k}$. Consider the natural action of $D_{4k}$ on the vertices of the $2k$-gon. We want to compute 
$\text{Fix}(g)$ for each conjugacy class.


Every rotation (excluding the identity) will fix no points. Similarly an odd number of rotations and then a flip 
will never yield any fixed points.  The only elements that will fix anything will be the identity (which will fix 
all $2k$ vertices) and the primitive reflections, which will each fix exactly the two vertices that lie in the axis 
of reflection. Thus the permutation character $\pi(g) = \text{fix}(g)$ is given by
\begin{center}
\begin{tabular}{c|ccccc}
    $g$ &$1$ & $a^k$ & $a^i$ & $b$ & $ab$ \\ \hline
    $\pi(g)$ & $2k$ & $0$ & $0$ & $2$ & $0$ \\
\end{tabular}
\end{center}
for all $1 \leq i \leq k-1$.

\subsection{Queries for Exact Learning}

We want to determine $\gamma(D_{4k})$. In order to do this, we look at the irreducible constituents of the tensor 
powers of the permutation character. We need to find the smallest integer $t$ such that $\pi^t$, the $t$th tensor 
power of pi, contains every irreducible character as a constituent. That is, the smallest $t$ such that $\langle 
\pi^t, \chi \rangle \neq 0$ for all irreducible characters $\chi$.


We summarize the information we have computed so far about the characters of $D_{4k}$

\def\cl{\text{cl}}
\begin{center}
\begin{tabular}{c|ccccc}
    $g$ &$1$ & $a^k$ & $a^i$ & $b$ & $ab$ \\ \hline
    $|\cl(g)|$ & $1$ & $1$ & $2$ & $k$ & $k$ \\ \hline
%
    $\lambda_1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
    $\lambda_2$ & $1$ & $1$ & $1$ & $-1$ & $-1$ \\
    $\lambda_3$ & $1$ & $(-1)^k$ & $(-1)^i$ & $1$ & $-1$ \\
    $\lambda_4$ & $1$ & $(-1)^k$ & $(-1)^i$ & $-1$ & $1$ \\
    $\psi_j$ & $2$ & $2(-1)^j$ & $\zeta_{2k}^{ij} + \zeta_{2k}^{-ij}$ & $0$ & $0$ \\ \hline
    $\pi$ & $2k$ & $0$ & $0$ & $2$ & $0$ 
\end{tabular}
\end{center}
where $\psi_j$ is a family of representations, one for each $1 \leq j \leq k-1$ and $a^i$ is a unique conjugacy 
classes for each $1 \leq i \leq k-1$.

To determine which irreducible characters are constituents of $\pi$, we compute $\langle \pi, \chi\rangle$ for each 
irreducible $\chi$.  Since $\pi$ is nonzero only on the classes $1$ and $b$ we omit these terms.
\begin{align*}
    \langle \pi, \lambda_1 \rangle &= \frac{1}{|G|}\bigg( (1)(1)(2k) + (2)(1)(k) \bigg)  = \frac{1}{4k}(2k + 2k) = 
    1 \\
%
    \langle \pi, \lambda_2 \rangle &= \frac{1}{|G|}\bigg( (1)(1)(2k) + (2)(-1)(k) \bigg)  = \frac{1}{4k}(2k - 2k) = 
    0 \\
%
    \langle \pi, \lambda_3 \rangle &= \frac{1}{|G|}\bigg( (1)(1)(2k) + (2)(1)(k) \bigg)  = \frac{1}{4k}(2k - 2k) = 
    1 \\
%
    \langle \pi, \lambda_4 \rangle &= \frac{1}{|G|}\bigg( (1)(1)(2k) + (2)(-1)(k) \bigg)  = \frac{1}{4k}(2k - 2k) = 
    0 \\
    \langle \pi, \psi_j \rangle &= \frac{1}{2n}(2n) = 1 
\end{align*}

Thus, the decomposition of $\pi$ into irreducibles is given by
\[
    \pi = \lambda_1 + \lambda_3 + \sum_{j=1}^{k-1} \psi_j
\]
Hence, the irreducibles $\lambda_2$ and $\lambda_4$ are not constituents of $\pi$ and so $\gamma(D_{4k})$ must be 
greater than one. By Theorem \ref{Burnside-Brauer}, every irreducible must appear in the decomposition of $\pi^t$ 
for some $t \leq 2$, thus we know without any further computation that $\gamma(D_{4k}) = 2$.



\subsection{Queries for bounded probability of learning}
Although this means a quantum oracle does not help you solve the problem exactly with fewer queries, it is better 
when you allow for finding a solution with high probability.

As given in chapter 3, the number of queries needed to succeed with probability greater than 2/3 is the smallest 
positive integer $t$ such that
\[
    \frac{1}{|G|} \sum_{\chi \in N^{\otimes t}} \chi(e)^2  \ \geq\  \frac{2}{3}
\]

We can find this value using what we computed in section 4.2.2.  The constituent irreducibles of $\pi$ are 
$\lambda_1, \lambda_3,$ and $\psi_j$ for each $1 \leq j \leq k-1$. These are 2 characters of degree 1 and $k-1$ 
characters of degree 2. Thus
\[
    \frac{1}{|D_{4k}|} \sum_{\chi \in N^{\otimes t}} \chi(e)^2 = \frac{1}{4k}\bigl( 2(1^2) + (k-1)(2^2) \bigr) = 
    \frac{1}{4k}(2 + 4k - 4) = 1 - \frac{1}{2k}
\]

Therefore one can succeed using only one query with probability $1 - \frac{1}{2k}$, which is bounded above 2/3 for 
every dihedral group of order $4k$.

\section{The General Linear Group GL(2,q)}

Here we present the analogous results for $\GL(2, q)$, the group of $2 \times 2$ invertible matrices over the 
finite field with $q$ elements. 



\subsection{Conjugacy Classes}

The conjugacy classes of $\GL(2,q)$ can be classified into four families.
\begin{enumerate}
    \item Scalar multiples of the identity i.e. matrices of the form
        \[
            \rmat{a}{0}{0}{a} 
        \]
        for some $a \in \F_q$. Denote this family of matrices $aI$.
    \item Upper Triangle 
\end{enumerate}









%# Move to chapter 3
%For a fixed group, this process is purely mechanical. For each irreducible character $\chi$, we compute the 
%projection of $\pi$ onto $\lambda$ by taking the inner-product $\langle \pi, \lambda \rangle$. If this 
%inner-product is zero for any $\chi$, then it is not a constituent of $\pi$. We repeat this process for each 
 


